---
title: "blackfridaysales"
author: "rishabhkansal"
date: "2023-01-19"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = '', 
                      warning = FALSE, 
                      message = FALSE)
```

```{r message=FALSE, warning=FALSE}
options(digits = 3, scipen = 999)
library(tidyverse)
library(ISLR)
library(jtools)
library(caret)
library(ROCR)
library(e1071)
library(glmnet)
library(lmtest)
library(rpart)
library(rpart.plot)
library(caTools)
set.seed(2)
rm(list=ls())
train <- read.csv("train.csv")
```

```{r}
col_names <- c("Gender","Age","Occupation","City_Category","Marital_Status","Stay_In_Current_City_Years")
train[col_names] <- lapply(train[col_names], factor)
```


```{r}
head(train)
```

```{r}
str(train)
```
```{r}
library(naniar)
vis_miss(train, warn_large_data = FALSE)
```
Product categories 2 and 3 have a fair amount of null values. The rest of the data set seems fine.
```{r}
ggplot(data = train, aes(x = Purchase)) + 
  geom_histogram(binwidth = 1000, color = "black", fill = "white") +
  geom_density(col = "red") +
  ggtitle("Distribution of Purchase") +
  xlab("Purchase") + 
  ylab("Frequency") +
  scale_fill_manual(values=c("#F5A623", "#F5A623", "#F5A623"))
```
Outlier Correction 

```{r}
boxplot(train$Purchase)
```
```{r}
#calculate the upper and lower limits for outlier detection
purchase_mean <- mean(train$Purchase)
purchase_sd <- sd(train$Purchase)
upper_limit <- purchase_mean + 3*purchase_sd
lower_limit <- purchase_mean - 3*purchase_sd

# Replace outliers with the mean of the column
train$Purchase[train$Purchase < lower_limit | train$Purchase > upper_limit] <- purchase_mean
```

#EDA
```{r}
library(gridExtra)
col_names <- c("Gender","Age","City_Category","Marital_Status")
# Initialize an empty list to store the plots
plots <- list()
# Iterate through the columns and create a pie chart for each one
for (col in col_names) {
    plots[[col]] <- ggplot(train, aes_string(x = col, fill = col)) + 
    geom_bar(width = 1) +
    coord_polar("y", start = 0) +
    ggtitle(paste0("Distribution of ", col)) +
    theme(legend.position = "none") +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
}
# Use grid.arrange to combine the plots into one figure
grid.arrange(grobs = plots, ncol = 3)
```
Data is fairly balanced in terms of Age and City but it seems to have excessive single males. 

#EDA
```{r}
col_names <- c("Gender","Age","City_Category","Marital_Status")
plots <- list()
for (col in col_names) {
    plots[[col]] <- ggplot(train, aes_string(x = col, y = "Purchase", fill = col)) +
    geom_bar(stat = "identity") +
    ggtitle(paste0("Cum Purch by ", col, " Category")) +
    xlab(col) +
    ylab("Cumulative Purchases")
}
grid.arrange(grobs = plots, ncol = 2)
```
Takeaways:
1. Male customers contribute to most of the sales. On average, males spent more money
2. 59% of customers are not married and 41% of customers are married. For spending, there are no any difference

#Average ticket analysis 
```{r}
col_names <- c("Gender","Age","City_Category","Marital_Status","Stay_In_Current_City_Years")
results <- data.frame()
for (col in col_names) {
    factors <- unique(train[,col])
    for (factor in factors) {
        filtered_data <- train[train[,col] == factor,]
        avg_purchase <- mean(filtered_data$Purchase)
        results <- rbind(results, data.frame(col = col, factor = factor, avg_purchase = avg_purchase))
    }
}
split_df <- split(results, results$col)
plots_list <- list()
for (name in names(split_df)) {
    df <- split_df[[name]]
    p <- ggplot(df, aes(x = factor, y = avg_purchase)) +
         geom_bar(stat = "identity", fill = avg_purchase, color = "black") +
         ggtitle(name) + 
         theme_minimal() + 
         theme(plot.title = element_text(hjust = 0.5))+
         ylab("Avg Purchase")+
         xlab("Factor") +
         theme(legend.position = "none")
    plots_list[[name]] <- p
}
library(gridExtra)
grid.arrange(grobs = plots_list, ncol = 3)

```
There is not much difference in the average buying behavior for each column. 
1. Males spend a little bit more than women on average, combined with a larger number of men in the sample, this means that they are responsible for a majority of total purch

#Making Predicitons using xGBoost

```{r}
library(xgboost)
train_data <- train[,!(names(train) %in% c("User_ID", "Product_ID","Occupation"))]
col_names = c("Gender","Age","City_Category","Marital_Status","Stay_In_Current_City_Years")
train_data <- as.data.frame(cbind(train_data,model.matrix(~.-1, train_data[col_names])))
train_data$Age <- as.numeric(as.character(train_data$Age))
set.seed(123)
split <- sample.split(train_data$Purchase, 0.7)
train_data <- subset(train_data, split == TRUE)
test_data <- subset(train_data, split == FALSE)
train_data <- train_data[,!(names(train_data) %in% c("Gender","Age","City_Category","Marital_Status","Stay_In_Current_City_Years"))]
test_data <- test_data[,!(names(test_data) %in% c("Gender","Age","City_Category","Marital_Status","Stay_In_Current_City_Years"))]
# Train the model
xgb_model <- xgboost(data = as.matrix(train_data[, -1]), label = train_data$Purchase, nrounds = 10)
# Make predictions on test data
predictions <- predict(xgb_model, as.matrix(test_data[, -1]))
# Create a data frame for actual and predicted values
pred_df <- data.frame(actual = test_data$Purchase, predicted = predictions)

ggplot() + 
  geom_density(aes(x=test_data$Purchase, fill='blue'),alpha = 0.5) + 
  geom_density(aes(x=predictions, fill='red'), alpha = 0.5) + 
  ggtitle("Actual vs Predicted Purchase") +
  xlab("Purchase") +
  ylab("Density")+
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(name = "", labels = c("Actual", "Predicted"), values = c("blue", "red"))+
  scale_x_continuous(limits=c(0,20000))
```
